### 2016-04-25 papers/asonam2016/paper.tex
More progress

\documentclass[conference]{IEEEtran}
\usepackage[style=ieee, backend=biber]{biblatex}

\ifCLASSINFOpdf
	\usepackage[pdftex]{graphicx}
	\graphicspath{{img/}}
	\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\else
	\usepackage[dvips]{graphicx}
 	\graphicspath{{img/}}
	\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\fi

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\bibliography{paper.bib}

\begin{document}
\title{Mimicry in Online Conversations: An Exploratory Study of Linguistic Analysis Techniques}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Tom Carrick, Awais Rashid, Paul J Taylor}
\IEEEauthorblockA{Security Lancaster / CREST\\
Lancaster University\\
Lancaster, United Kingdom\\
Email: \{t.carrick, a.rashid, p.j.taylor\} @lancaster.ac.uk}
\and
\IEEEauthorblockN{Adam Joinson}
\IEEEauthorblockA{???\\
University of Bath\\
Bath, United Kingdom\\
Email: a.joinson@bath.ac.uk}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Interactive Alignment Model is the prevailing theory of
[I probably need some help phrasing this].
In this study we investigate how several linguistic alignment techniques can be used
to detect mimicry at the lexical, syntactic and situation levels of the
Interactive Alignment Model.

We show that [something].
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}

Mimicry in online conversations is an important and active area of research. Humans tend to mimic each
other in many areas, such as speech
\cite{levelt1982surface, bock1986syntactic, bock1989closed, garrod1987saying},
physical actions\cite{bernieri1988coordinated}
and written text{\cite{danescu2011chameleons, scissors2008linguistic}.
People that mimic each other's language have more successful outcomes in joint tasks
\cite{maddux2008chameleons}. Additionally, mimicry can be an indicator of insider threats
\cite{taylor2013detecting}.

There are a number of measures designed to capture mimicry in text, and some efforts have been made to
compare and trial these techniques. However, few of these techniques are grounded what cognitive 
science knows about how mimicry works within humans.

This perspective is important so that each measure is grounded in reality.
Moreover, it suggests that the best approach may be to combine existing measures to create
a more holistic technique for mimicry detection.

The most similar piece of work is a comparison of several methods\cite{xu2015evaluation}, some 
discussed in this paper. However, it does not try to directly compare the accuracy of the measures,
nor does it look at how they fit into the Interactive Alignment Model.

This paper aims to compare how well existing linguistic analysis techniques work to detect mimicry in 
relation to the Interactive Alignment Model. Further, we also look at some existing techniques that 
have not previously been used to detect mimicry and discuss their viability.

The novel contributions of this paper are:
\begin{itemize}
	\item Compares existing mimicry detection techniques and how they map to the
		  Interactive Alignment Model
	\item Trials techniques that have not been used for mimicry detection before.
	\item Looks at creating hybrid techniques, combining measures together to find if they are more 
		  accurate when combined. 
\end{itemize}

This paper is divided into six sections. The next section introduces the
Interactive Alignment Model, the psychological model of mimicry we are testing our techniques against.
Section III gives a background of each measure tested in the paper, and defines which layer of the
IAM (if any) the measure falls into.
Section IV addresses the data collected and how it is processed for each technique.
We follow in section V with a discussion of the results, and finally give our conclusions and
suggestions for further work in section VI.

\section{The Interactive Alignment Model}
The Interactive Alignment Model (IAM)\cite{pickering2004toward} is an attempt to model alignment in
conversations. It consists of six layers, which with the exception of the top layer, the situation
model, are all representations of some aspect of speech.

\begin{enumerate}
	\item The situation model --- a representation of the situation being discussed
	\item The semantic representation --- the meaning of the utterance
	\item The syntactic representation --- the sentence structure
	\item The lexical representation --- the choice of words used
	\item The phonological representation --- ???
	\item The phonetic representation --- the sounds of an utterance
\end{enumerate}

The phonological and phonetic representations are not discussed in this paper as they relate purely
to the sound of speech, representations that do not exist in the written text of online conversations.

Further, the situation model is not included as there are currently no measures that can be used for
finding alignment at the situational level.

The theory states that alignment is automatic, that conversation partners will align their speech 
towards each other at every level of the model. In addition, alignment "percolates" between layers such
that increased alignment in one layer will lead to increased alignment in other layers.

\begin{figure*}
	\caption{The interactive alignment model}
	\includegraphics[width=\textwidth]{iam}
\end{figure*}

\section{Analysis Techniques}

\subsection{Local Linguistic Alignment}
Local Linguistic Alignment (LLA) is a simple,  dual purpose method that can work at both the lexical
and syntactic layers\cite{wang2014linguistic, xu2015evaluation}.

Lexical Indiscriminate Local Linguistic Alignment (LILLA) works at the lexical layer. It gives the
probability that a word from the prime document will appear at a particular position in the target 
document. It is defined as:

\[LILLA(target, prime) = \frac{p(target|prime)}{p(target)}\]

Syntactic Indiscriminate Local Linguistic Alignemnt (SILLA) works at the syntactic layer. It works
in exactly the same way as LILLA, but rather than compare words, each sentence is annotated with a
phrase structure tree, and LLA is run on these instead. As such, it measures the probability of a
phrase structure subtree in the prime document appearing occurring at a given position in the target
document.

\subsection{Hierarchical Alignment Model}
\cite{doyle2016robust}

\subsection{Word Vectors}
word2vec\cite{mikolov2013distributed}
doc2vec\cite{le2014distributed}.

\subsection{Linguistic Style Matching}
\cite{niederhoffer2002linguistic}

\section{Method}
The dataset consists of 103 dyadic conversations taken from Twitter. Tweets were found via Twitter's 
search API, using English language and being a reply as criteria. We then work up the chain of tweets 
until we hit the end, or a third interlocutor is found, at which point the conversation is cut off and 
kept, as long as there are no fewer than 10 tweets in the conversation. 200 tweets were found using 
this method. Of these, 62 were removed as they were not true conversations (either monologues or 
collaborative fiction writing/roleplay). 18 conversations were removed as they had significant amounts 
of non-English text, and 17 conversations involved bots and were also removed.

The data is processed as follows:
\begin{itemize}
	\item Screen names at the start of the message are removed. These are just the names of the 
		  interlocutors and only add noise.
	\item All URLs are replace with the string '[URL]'.
	\item HTML entities such as \&amp; are converted to plain text (\&).
	\item Any adjacent messages by the same author are merged together such that each message always
	      alternates between the two interlocutors.
\end{itemize}

[LIMITATIONS]

The data was classified by three researchers with backgrounds in mimicry independent of the authors of 
this paper.
Each  conversation was classified for lexical, syntactic and semantic mimicry on a three point scale 
where 0 represents no mimicry, 1 represents low levels of mimicry at that layer, and 2 represents high 
levels of mimicry. Classifications from each author were averaged together to give each conversation
three scores between 0 and 2.

Each measure is then run against the data such that each message from author A is compared with each 
message from author B that occurred before the original message. The exception for this is HAM,
where we have used the original implementation [need to actually do this and explain it further].
\footnote{[I could probably make my github repo public and put a link to the code here?]}

\section{Results}
Still not sure how to compare everything.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Discussion}
Discussion

\section{Conclusion and Future Work}
The conclusion goes here.
Zelig\cite{jones2014finding}

% use section* for acknowledgment
\section*{Acknowledgment}
The authors gratefully acknowledge a grant from CREST (probably need to find the name of this).

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

\printbibliography

\end{document}